The influence of habitat quality and environmental transmission on host-pathogen metapopulation dynamics

### 2014-10-16 Manuscript/manuscript.tex
Notational changes
Changing notation so that all parameters are greek letters.  Avoids
confusion between e and exponential, and between E and empty state; also
tweaked so gamma is a function of tau, the time since infection.

\begin{table}
\caption{State transitions and their rates for patch $i$ in a metapopulation simulation.  $S$ denotes occupied by  susceptible hosts, $I$ denotes occupied infectious hosts, and $\emptyset$ denotes unoccupied by the host.  All patches are characterized by an environmental infection rate, $\gamma_i$, which is non-zero only if the patch has been previously occupied by infectious hosts.}
\begin{tabular}{lr}
State Transition & Rate \\
\hline
$S \rightarrow I$ & $\delta C_{Ii} + \gamma_i$\\
$S \rightarrow \emptyset $ & $e_{Si}$\\
$I \rightarrow \emptyset $ & $e_{Ii}$ \\
$\emptyset \rightarrow S$ & $C_{Si}$\\
$\emptyset \rightarrow I$ & $C_{Ii}$\\
\label{transitions}
\end{tabular}
\end{table}

@@

Building on the framework developed by \cite{Hanski2003}, the rate at which patch \emph{i}, with quality $A_i$, is colonized by individuals in state $X \in (S,I)$ is given by its connectivity:
\begin{equation}
C_{Xi}=A_i^{\xi_{im}} \sum_{i\neq j }\phi_jA_j^{\xi_{em}}e^{-D d_{ij}},
\label{connectivity}
\end{equation}
where $\phi_j$ is an indicator function that is $1$ if patch $j$ is in state $X$ and is $0$ otherwise; $\xi_{im}$ and $\xi_{em}$ control how rates of immigration and emigration, respectively, scale with patch quality; $D$ is the inverse of mean dispersal distance; and $d_{ij}$ is the distance between patches $i$ and $j$.  Essentially, $C_{Xi}$ sums the colonization effort from all patches in state $X$ to patch $i$.

The extinction rate of patch $i$ in state $X$ is given by:
\begin{equation}
E_{Xi}=\frac{e_X}{A_i^\alpha},
\end{equation}
where $e_X$ is the extinction rate of a patch of unit quality in state $X$ (here we assume that $e_I \geq e_S$ to account for disease-induced mortality), and $\alpha$ controls how extinction rate scales with patch quality.  When an infectious population goes extinct, we assume that the hosts leave behind an infectious pathogen reservoir on the patch.

Susceptible populations are infected via direct contact with infectious colonists at rate $\delta C_{Ii}$, where $\delta$ is the direct transmission probability.  In addition, environmental transmission can take place when a susceptible population occupies a previously infected patch, and occurs at rate:
\begin{equation}
\gamma_i=\gamma_0exp(-rt_{Ii}),
\end{equation}
where $\gamma_0$ is the initial infection rate of the pathogen reservoir, $t_{Ii}$ is the time since last infectious occupancy ($t_{Ii} = \infty$, and thus $\gamma_i = 0$, if the patch has never been occupied by infectious hosts), and $r$ is the pathogen's decay rate in the environment.

\begin{table}[h!]   
\caption{Parameters of the SPOM model, their meaning, and the values assigned under different parameterizations.  Empty cells indicate the same value as the default parameterization.}
\begin{tabular}{l p{5cm} rrrr}
Parameter & Interpretation &  Default & Lattice & Equal Ext. & Weak Direct Inf.\\
\hline
$\xi_{im}$ & Scaling parameter for effect of target patch quality on immigration & 0.5 & -- & -- & --\\
$\xi_{em}$ & Scaling parameter for effect of target patch quality on emigration & 0.5 & -- & -- & --\\
$D$& Inverse of mean dispersal distance & 5 & 2 & -- & --\\
$d_{ij}$ & Distance between patch $i$ and $j$ & 1 $\forall i \neq j$ & 1 $\forall i, j$ neighbors & -- & --\\
\hline
$e_S$ & Extinction rate of unit quality susceptible patch & 0.1 & -- & -- & --\\
$e_I$ & Extinction rate of unity quality infectious patch & 0.5 & -- & -- & --\\
$\alpha$ & Strength of environmental stochasticity & 1 & -- & 0 & --\\
\hline
$\delta$ & Probability of direct infection & 0.5 & -- & -- & 0.1\\
$\gamma_0$ & Initial rate of infection from reservoir patch & 0.5 & -- & -- & --
\end{tabular}
\label{params}
\end{table}

>>>>

\begin{table}
\caption{State transitions and their rates for patch $i$ in a metapopulation simulation.  $S$ denotes occupied by  susceptible hosts, $I$ denotes occupied infectious hosts, and $\emptyset$ denotes unoccupied by the host.  All patches are characterized by an environmental infection rate, $\gamma(\tau_i)$, which is non-zero only if the patch has been previously occupied by infectious hosts.}
\begin{tabular}{llr}
State Transition & Process &  Rate \\
\hline
$S \rightarrow I$ & Infection (from contact or reservoir) & $\delta \lambda_{Ii} + \gamma(\tau_i)$\\
$S \rightarrow \emptyset $ & Extinction (of susceptible) & $\epsilon_{Si}$\\
$I \rightarrow \emptyset $ & Extinction (of infectious) &  $\epsilon_{Ii}$ \\
$\emptyset \rightarrow S$ & Colonization (by susceptibles) & $\lambda_{Si}$\\
$\emptyset \rightarrow I$ & Colonization (by infectious) & $\lambda_{Ii}$\\
\label{transitions}
\end{tabular}
\end{table}

@@

Building on the framework developed by \cite{Hanski2003}, the rate at which patch \emph{i}, with quality $A_i$, is colonized by individuals in state $X \in (S,I)$ is given by its connectivity:
\begin{equation}
\lambda_{Xi}=A_i^{\xi_{im}} \sum_{i\neq j }\phi_jA_j^{\xi_{em}}\exp(-D d_{ij}),
\label{connectivity}
\end{equation}
where $\phi_j$ is an indicator function that is $1$ if patch $j$ is in state $X$ and is $0$ otherwise; $\xi_{im}$ and $\xi_{em}$ control how rates of immigration and emigration, respectively, scale with patch quality; $D$ is the inverse of mean dispersal distance; and $d_{ij}$ is the distance between patches $i$ and $j$.  Essentially, $C_{Xi}$ sums the colonization effort from all patches in state $X$ to patch $i$.

The extinction rate of patch $i$ in state $X$ is given by:
\begin{equation}
\epsilon_{Xi}=\frac{\mu_X}{A_i^\alpha},
\end{equation}
where $\mu_X$ is the extinction rate of a patch of unit quality in state $X$ (here we assume that $\mu_I \geq \mu_S$ to account for disease-induced mortality), and $\alpha$ controls how extinction rate scales with patch quality.  When an infectious population goes extinct, we assume that the hosts leave behind an infectious pathogen reservoir on the patch.

Susceptible populations are infected via direct contact with infectious colonists at rate $\delta \lambda_{Ii}$, where $\delta$ is the direct transmission probability.  In addition, environmental transmission can take place when a susceptible population occupies a previously infected patch, and occurs at rate:
\begin{equation}
\gamma(\tau_i)=\gamma_0exp(-r\tau_{i}),
\end{equation}
where $\gamma_0$ is the initial infection rate of the pathogen reservoir, $\tau_{i}$ is the time since last infectious occupancy ($\tau_{i} = \infty$, and thus $\gamma(\tau_i) = 0$, if the patch has never been occupied by infectious hosts), and $r$ is the pathogen's decay rate in the environment.

\begin{table}[h!]   
\caption{Parameters of the SPOM model, their meaning, and the values assigned under different parameterizations.  Empty cells indicate the same value as the default parameterization.}
\begin{tabular}{l p{5cm} rrrr}
Parameter & Interpretation &  Default & Lattice & Equal Ext. & Weak Direct Inf.\\
\hline
$\xi_{im}$ & Scaling parameter for effect of target patch quality on immigration & 0.5 & -- & -- & --\\
$\xi_{em}$ & Scaling parameter for effect of target patch quality on emigration & 0.5 & -- & -- & --\\
$D$& Inverse of mean dispersal distance & 5 & 2 & -- & --\\
$d_{ij}$ & Distance between patch $i$ and $j$ & 1 $\forall i \neq j$ & 1 $\forall i, j$ neighbors & -- & --\\
\hline
$\mu_S$ & Extinction rate of unit quality susceptible patch & 0.1 & -- & -- & --\\
$\mu_I$ & Extinction rate of unity quality infectious patch & 0.5 & -- & -- & --\\
$\alpha$ & Strength of environmental stochasticity & 1 & -- & 0 & --\\
\hline
$\delta$ & Probability of direct infection & 0.5 & -- & -- & 0.1\\
$\gamma_0$ & Initial rate of infection from reservoir patch & 0.5 & -- & -- & --
\end{tabular}
\label{params}
\end{table}
